### 2024-03-02 paper/paper.tex
fixed issues of CNN

\documentclass{article}
\usepackage{geometry}
\usepackage{cite}
\geometry{a4paper}

\title{Text-Mining Short Paper}
\author{Qixuan Yang, Matei-Alexandru Costin, Hong Wang, Jianxin YU}
\date{2nd March 2024}

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
% \tableofcontents

\section{Introduction}
\cite{yao2019docred}

\section{Method}

\bibliography{references}
\bibliographystyle{plainnat}

\end{document}